You are the way you structurally talk: structural-temporal neighbourhoods of posts to characterize users in online forums

### 2016-04-23 doc/IEEEtran/paper_IEEEformat.tex
paper ++

\subsection{Time-based}
In a first attempt to take time into account, one might set fixed time-based boundaries for the neighbourhood and include only those posts whose timestamp $t_j$ is at distance less than $\tau$ from the ego post $i$, $|t_j-t_i|<\tau$. However, the pace at which posts are added to the conversation may be very different between conversation threads (and also within a thread) and we have no \textit{a priori} criteria for a proper choice of $\tau$. 
Rather than looking for a fixed time radius, we may look at changes in the pace how new posts are added to the thread. In statistical analysis, a \textit{changepoint} in a sequence $x_1,...x_n$ is a point that comes from a different probability distribution than its precedent values. If applied either to the timestamps of posts in a branch (vertical sequence) or to the timestamps of the replies to the same post (horizontal sequence), we say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$. Now we can introduce our second definition of neighbourhood.
\begin{definition}
Given a tree graph $G$, the \textit{time-based neighbourhood} of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic than $i$.
\end{definition}
An example of time-based neighbourhood is given in Figure~\ref{fig:cutpoints} (right).
Note that we still have a radius parameter $r$ to guarantee that the resulting structure remains local even if no changepoint is detected near the post $i$. Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods according to the given definition. 
\begin{algorithm}[H]
\begin{algorithmic}
\REQUIRE Posts tree $g$, vertical breakpoints, horizontal breakpoints, ego post $ego$
\ENSURE Subgraph of $g$ with all vertices in $V(g)$
\STATE Compute structural neighbourhood $\mathcal{N}_i(r)$
\STATE ancestors $\leftarrow$  ancestors(ego) in $\mathcal{N}_i(r)$
\STATE older\_siblings $\leftarrow$ older\_siblings(ego) in $\mathcal{N}_i(r)$
\STATE dump $\leftarrow \varnothing$
\FOR{bp $\in$ vertical breakpoints}
 \IF{bp $\in$ ancestors}
   \STATE dump $\leftarrow$ dump $\cup$ ancestors(bp)
 \ELSE
   \STATE dump $\leftarrow$ dump $\cup$ descendants(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\FOR{bp $\in$ horizontal breakpoints}
  \IF{bp $\in$ (older\_siblings $\cup$ ancestors)}
     \STATE dump $\leftarrow$ dump $\cup$ older\_siblings(bp)
   \ELSE
     \STATE dump $\leftarrow$ dump $\cup$ younger\_siblings(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\STATE $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts) from $\mathcal{N}_i(r)$
\end{algorithmic}
\caption{Extraction of time-based neighbourhood}
\label{alg:temporal_neighbourhood}
\end{algorithm}

@@

\section{Application to Reddit forums}
We applied our two definitions of structural-temporal neighbourhood (order-based and time-based) to a forum dataset. Our dataset consists of all posts from March 2014 to May 2015 of the Podemos forum in Reddit\footnote{https://www.reddit.com/r/podemos}. The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid the October 18th and 19th 2014. Nowadays, its members use it mainly to share and discuss about political news.
Our dataset contains contains 75,000 posts spread over 4,262 threads and written by 9,160 users between the April 25th and October 20th 2014.
\subsection{Neighbourhood extraction}
As a previous step for all further analysis we detect the neighbourhood around every post in our datasets. Each neighbourhood structure (including colours) is given a unique integer number. Unfortunately, it is not practical to give a number to every theoretically possible structure since, even if we could enumerate all the possible trees up to a given size, a lot of of these trees would not be seen in the data. It appears then a much more practical approach to enumerate only those trees that are seen in the data. The consequence of this is that every dataset will have its own dictionary of trees.

@@

\subsection{Agreement between neighbourhood methods}
The detection of the time-based and order-based neighbourhood in the Game of Thrones dataset gave us a list of 2000 different (non-isomorphic) time-based neighbourhoods and 180 different (non-isomorphic) \footnote{TODO: Review numbers} order-based neighbourhoods, which means that the time-based is able to capture a much wider set of structures. Of course, this is not necessarily and advantage since these extra structures may be just spurious and only occur a very small number of times. Indeed, Figure \ref{fig:census_distributions} shows that around half of the time-based neighbourhoods occur only once, and that we only need 384 time-based to account for the 95\% of the census, while for the order-based, 53 structures are enough to cover the same percentage. Moreover, there is no much difference, in terms of frequency, between the 50 most frequent neighbours of each type. The reason why the frequency of the order-based neighbourhoods decay faster might be that the most frequent neighbourhood is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.

@@

In Figure \ref{fig:census_compare} we compare the neighbourhood census in two different forums, Podemos (politics, news and discussion among party supporters) and Game of Thrones (discussion around the TV series). Interestingly, two time-based neighbourhoods are  more than three times likelier in the Game of Thrones. These neighbourhoods correspond to a discussion between two users (21) and to XXXXXX (25). The same comparison for the order-based triad is harder to interpret, since the outstanding neighbourhoods in the Game of Thrones correspond to XXXX (25) and ... (27)

@@

The analysis followed in this section lead us to think that the time-based neighbourhood is more convenient than the order-based. The main reason, we think, is that order seems to be a too dogmatic way to decide where a neighbourhood ends.

>>>>

\subsection{Time-based}
In a first attempt to take time into account, one might set fixed time-based boundaries for the neighbourhood and include only those posts whose timestamp $t_j$ is at distance less than $\tau$ from the ego post $i$, $|t_j-t_i|<\tau$. However, the pace at which posts are added to the conversation may be very different between conversation threads (and also within a thread) and we have no \textit{a priori} criteria for a proper choice of $\tau$. 
Rather than looking for a fixed time radius, we may look at changes in the pace how new posts are added to the thread. In statistical analysis, a \textit{changepoint} in a sequence $x_1,...x_n$ is a point that comes from a different probability distribution than its precedent values. If applied either to the timestamps of posts in a branch (vertical sequence) or to the timestamps of the replies to the same post (horizontal sequence), we say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$. Now we can introduce our second definition of neighbourhood.
\begin{definition}
Given a tree graph $G$, the \textit{time-based neighbourhood} of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic than $i$.
\end{definition}
For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by its own authors is available in the R library \texttt{changepoint}.}. An example of time-based neighbourhood is given in Figure~\ref{fig:cutpoints} (right).
Note that we still have a radius parameter $r$ to guarantee that the resulting structure remains local even if no changepoint is detected near the post $i$. Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods according to the given definition. 
\begin{algorithm}[H]
\begin{algorithmic}
\REQUIRE Posts tree $g$, vertical breakpoints, horizontal breakpoints, ego post $ego$
\ENSURE Subgraph of $g$ with all vertices in $V(g)$
\STATE Compute structural neighbourhood $\mathcal{N}_i(r)$
\STATE ancestors $\leftarrow$  ancestors(ego) in $\mathcal{N}_i(r)$
\STATE older\_siblings $\leftarrow$ older\_siblings(ego) in $\mathcal{N}_i(r)$
\STATE dump $\leftarrow \varnothing$
\FOR{bp $\in$ vertical breakpoints}
 \IF{bp $\in$ ancestors}
   \STATE dump $\leftarrow$ dump $\cup$ ancestors(bp)
 \ELSE
   \STATE dump $\leftarrow$ dump $\cup$ descendants(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\FOR{bp $\in$ horizontal breakpoints}
  \IF{bp $\in$ (older\_siblings $\cup$ ancestors)}
     \STATE dump $\leftarrow$ dump $\cup$ older\_siblings(bp)
   \ELSE
     \STATE dump $\leftarrow$ dump $\cup$ younger\_siblings(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\STATE $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts) from $\mathcal{N}_i(r)$
\end{algorithmic}
\caption{Extraction of time-based neighbourhood}
\label{alg:temporal_neighbourhood}
\end{algorithm}

@@

\section{Application to Reddit forums}
Our data consists of two different forums of Reddit, namely the Podemos dataset \footnote{https://www.reddit.com/r/podemos} and the Game of Thrones dataset\footnote{https://www.reddit.com/r/gameofthrones}. The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid the October 18th and 19th 2014. Nowadays, its members use it mainly to share and discuss about political news. The Game of Thrones is a casual discussion forum about the Game of Thrones TV series.
The Podemos dataset contains 75,000 posts corresponding to 4,262 threads written by 9,160 users between the April 25th and October 20th 2014. The Game of Thrones dataset contains 75,000 posts corresponding to XXX threads written by XXXXX users.

\subsection{Neighbourhood extraction}
As a previous step for all further analysis we detect the neighbourhood around every post in our datasets. Each neighbourhood structure (including colours) is given a unique integer number. Unfortunately, it is not practical to give a number to every theoretically possible structure since, even if we could enumerate all the possible trees up to a given size, a lot of of these trees would not be seen in the data. It appears then a much more practical approach to enumerate only those trees that are seen in the data. The consequence of this is that every dataset will have its own dictionary of trees.

@@

\subsection{Agreement between neighbourhood methods}
The detection of the time-based and order-based neighbourhood in the Game of Thrones dataset gave us a list of 2000 different (non-isomorphic) time-based neighbourhoods and 180 different (non-isomorphic) \footnote{TODO: Review numbers} order-based neighbourhoods, which means that the time-based is able to capture a much wider set of structures. Of course, this is not necessarily and advantage since these extra structures may be just spurious and only occur a very small number of times. Indeed, Figure~\ref{fig:census_distributions} shows that around half of the time-based neighbourhoods occur only once, and that we only need 384 time-based to cover 95\% of the census, while for the order-based, 53 structures are enough to cover the same percentage. Moreover, there is no much difference, in terms of frequency, between the 50 most frequent neighbours of each type. The reason why the frequency of the order-based neighbourhoods decay faster might be that the most frequent neighbourhood is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.

@@

In Figure \ref{fig:census_compare} we compare the neighbourhood census in the Podemos and the Game of Thrones forums. Interestingly, two time-based neighbourhoods are  more than three times likelier in the Game of Thrones. These neighbourhoods correspond to a discussion between two users (21) and to XXXXXX (25). The same comparison for the order-based triad is harder to interpret, since the outstanding neighbourhoods in the Game of Thrones correspond to XXXX (25) and ... (27)

@@

This results lead us to think that the time-based neighbourhood is more convenient than the order-based. The main reason, we think, is that order seems to be a too dogmatic way to decide where a neighbourhood ends.
